---
title: "Regularized and Sparse Generalized Canonical Correlation Analysis : user manual"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
   fig_align: 'center'
   fig_width: 6
   fig_height: 5
vignette: >
  %\VignetteIndexEntry{Regularized and Sparse Generalized Canonical Correlation Analysis : user manual}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, echo = FALSE, warning = FALSE, message = FALSE}

# Importing libraries
librairies = c("RGCCA", "ggplot2", "scales", "rgccaLauncher", "knitr", "pander")
for (l in librairies) {
  if (!(l %in% installed.packages()[, "Package"]))
    install.packages(l, repos = "http://cran.us.r-project.org", quiet = T)
  library(l, character.only = TRUE)
}


# Load the data
data("Russett")

#Creates the blocks
agriculture = Russett[, 1:3]
industry = Russett[, 4:5]
politic = Russett[, 6:11]

# Creates optional files
response = factor( apply(Russett[, 9:11], 1, which.max),
                   labels = colnames(Russett)[9:11] )
connection = matrix(c(0, 0, 0, 1,
                      0, 0, 0, 1,
                      0, 0, 0, 1,
                      1, 1, 1, 0),
                    4, 4)

# Save the Russett files
files = c("agriculture", "industry", "politic", "response", "connection")

# Creates the files (in .tsv format for example)
if(!file.exists("data"))
  dir.create("data")

# Without row and column names for connection files
sapply(1:length(files), function (x) {

  bool = ( files[x] != "connection")

  write.table(
    x = get(files[x]),
    file = paste("data/", files[x], ".tsv", sep=""),
    row.names = bool,
    col.names = bool,
    sep = "\t")
})

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "images/"
)

```


## 1.  Load the inputs
### Load the blocks

The blocks are loaded with the function ```setBlocks```. The first argument of this function (```superblock```) required a bolean giving the presence (TRUE) / absence (FALSE) of a superblock. The second one corresponds to a character giving the list of the file path separated by a comma (argument ```file```). By default, the name of the blocks correspond to those of the files (```names``` argument) adn could be set. By default, the tabulation is used as a column separator (```sep``` argument) and the first row is considered as a header (```header``` parameter).

```{r blocks}
# A boolean giving the presence (TRUE) / absence (FALSE) of a superblock
SUPERBLOCK =  TRUE

blocks = setBlocks(superblock = SUPERBLOCK, 
                   file = "data/agriculture.tsv,data/industry.tsv,data/politic.tsv")
```

### Load the groups of response and the connection between blocks

The connection between the blocks will be used by the RGCCA and must be set by ```setConnection``` function. A group of samples will be used to color them in the samples plot and must be set by ```setResponse``` function. For both functions, the ```blocks``` parameter, set at the previous step, is required. The other parameters are optionals. The user could import a file containing either (```file``` parameter) : (i) a symmetric matrix with 1 giving a connection between two blocs, or 0 otherwise; (ii) a univariate vector (qualtitative or quantitative) or a disjonctive table for the response. By default, the column separator is the tabulation and could be set (```sep``` argument). For the ```setResponse```, a header could be specified (```header``` parameter).

```{r connection}
# Optional parameters
RESPONSE = "data/response.tsv"
CONNECTION = "data/connection.tsv"
# uncomment the parameters below to try without default settings
# RESPONSE <- CONNECTION <- NULL

response = setResponse(blocks = blocks, 
                       file = RESPONSE)
connection = setConnection(blocks = blocks, 
                           file = CONNECTION)
```


### View the inputs

```{r views, echo=FALSE, results = "asis"}

for (x in 1:length(files)) {
  
  tab = get(files[x])
  if (x > 3) colnames(tab) = rep(NULL, NCOL(tab))
  
  pander::pandoc.table(head(tab),
                       caption = files[x])
}
```


## 2.  Run S/RGCCA

SGCCA is run from the RGCCA package by using two component for a biplot vizualisation. The S/RGCCA function doesn't names the blocks in their outputs. This step is required to generate biplots.

```{r rgcca}
# Use two component in RGCCA to plot in a bidimensionnal space
NB_COMP = c(2, 2, 3, 3)

sgcca.res = sgcca(A = blocks,
                  C = connection,
                  ncomp = NB_COMP)

# Renames the elements of the list according to the block names
names(sgcca.res$a) = names(blocks)
```


## 3.  Vizualise the analysis

Both the samples and the variables could be vizualized by using biplots functions (respectively ```plotSamplesSpace``` and ```plotVariablesSpace```). Histograms are used to vizualized in decreasing order the variables with the higher weights and the blocks with the higher Average Variance Explained (AVE).

These functions take the results of a sgcca or a rgcca (```rgcca``` parameter) and the components to vizualize : either ```comp_x``` and ```comp_y``` for biplots or  ```comp``` for histograms. By default, ```comp_x``` = ```comp``` = 1 and ```comp_y``` = 2. The presence or the absence of a superblock among the analysis could be specified for ```plotVariablesSpace``` and ```plotFingerprint``` to color the variables according to their blocks. By default, the last block is plotted, corresponding to the superblock if selected (```i_block``` parameter). ```plotVariablesSpace```, which is a corcircle plot, required the ```blocks``` for the correlation with the selected component. ```plotVariablesSpace``` could use the response variable to color the samples by groups. By default, the first 100th higher weights are used for the ```plotFingerprint``` and could be set by using the ```n_mark``` argument.

```{r}
COMP1 = 2
COMP2 = 3
NB_MARK = 100
```

### With the superblock, by default on the first and the second components

##### Samples plot
```{r}
plotSamplesSpace(rgcca = sgcca.res, 
                 resp = response)
```

##### Corcircle plot
```{r}
plotVariablesSpace(rgcca = sgcca.res,
                   block = blocks,
                   superblock = SUPERBLOCK)
```

##### Fingerprint plot
```{r}
plotFingerprint(rgcca = sgcca.res,
                superblock = SUPERBLOCK,
                n_mark = NB_MARK)
```

##### Best explained blocks
```{r}
plotAVE(rgcca = sgcca.res)

```

### With the politic block, on the 2nd and the 3rd components

##### Samples plot
```{r}
plotSamplesSpace(rgcca = sgcca.res, 
                 resp = response, 
                 comp_x = COMP1, 
                 comp_y = COMP2, 
                 i_block = 3)
```

##### Corcircle plot
```{r}
plotVariablesSpace(rgcca = sgcca.res,
                   block = blocks,
                   comp_x = COMP1,
                   comp_y = COMP2,
                   i_block = 3)
```

##### Fingerprint plot
```{r}
plotFingerprint(rgcca = sgcca.res,
                comp = COMP1,
                n_mark = NB_MARK, 
                i_block = 3)

```


```{r, echo = FALSE}
# Remove the temp/ folder
unlink("data", recursive = TRUE)
```
